---
title: "Ex1"
format: html
editor: visual
---

## Analytics of road closure and flooding

## active all the packages required

```{r,message=FALSE}
library(dplyr)
library(readr)
library(tidyverse)
```

## read all .csv files and bind rows

```{r, warning=FALSE}
Bega<- read_csv("data/Ex1/Bega.csv", skip = 7, show_col_types = FALSE, col_types = cols(.default = col_character()))
Cessnock<- read_csv("data/Ex1/Cessnock.csv", skip = 7, show_col_types = FALSE, col_types = cols(.default = col_character()))
Coffs_Harbour<- read_csv("data/Ex1/Coffs_Harbour.csv", skip = 7, show_col_types = FALSE, col_types = cols(.default = col_character()))
Dungog<- read_csv("data/Ex1/Dungog.csv", skip = 7, show_col_types = FALSE, col_types = cols(.default = col_character()))
Grafton<- read_csv("data/Ex1/Grafton.csv", skip = 7, show_col_types = FALSE, col_types = cols(.default = col_character()))
Kempsey<- read_csv("data/Ex1/Kempsey.csv", skip = 7, show_col_types = FALSE, col_types = cols(.default = col_character()))
Macksville<- read_csv("data/Ex1/Macksville.csv", skip = 7, show_col_types = FALSE, col_types = cols(.default = col_character()))
Maitland<- read_csv("data/Ex1/Maitland.csv", skip = 7, show_col_types = FALSE, col_types = cols(.default = col_character()))
Mid_coast<- read_csv("data/Ex1/Mid_coast.csv", skip = 7, show_col_types = FALSE, col_types = cols(.default = col_character()))
Newcastle<- read_csv("data/Ex1/Newcastle.csv", skip = 7, show_col_types = FALSE, col_types = cols(.default = col_character()))
Picton<- read_csv("data/Ex1/Picton.csv", skip = 7, show_col_types = FALSE, col_types = cols(.default = col_character()))
Port_Macquarie<- read_csv("data/Ex1/Port_Macquarie.csv", skip = 7, show_col_types = FALSE, col_types = cols(.default = col_character()))
Singleton<- read_csv("data/Ex1/Singleton.csv", skip = 7, show_col_types = FALSE, col_types = cols(.default = col_character()))

merged_data <- bind_rows(Bega, Cessnock, Coffs_Harbour, Dungog, Grafton, Kempsey, Macksville, Maitland, Mid_coast, Newcastle, Picton, Port_Macquarie, Singleton) |> 
  select(ID, Title, Location, Created, Ended, Version, Coordinates, Impact)
```

## tidy NA value of coordination, title, ID.

```{r}
road_coordination_tidied <- merged_data |> 
  filter(!is.na(Coordinates), Coordinates !="", Coordinates !="0") |> 
  filter(!is.na(ID), ID !="", ID !="0") |> 
  filter(!is.na(Title), Title !="", Title !="0")
```

## filt event caused by disaster

use filter() to keep the data which caused by natural disaster and delete repeat data. then write tidied data into new csv in the folder of "tidied data".

```{r,echo=FALSE}
disaster_stricken_road <-  road_coordination_tidied |>
  filter(Title %in% c("ADVERSE WEATHER Water over road", "FLOODING")) |> 
  filter(str_detect(Impact,"closed|Closed")) |> 
  distinct(ID, .keep_all = TRUE)
write_csv(disaster_stricken_road, "Tidied_data/Ex1/disaster_striken_road.csv")
```

# data descreption

## road closure

the data of road closure were downloaded in the website of [live Traffic NSW](https://www.livetraffic.com/historical-data-search), in the **Histroy** select the time from 15 May 2025 to 17 Jul 2025 with radius of 50km.

## meta data and relevant variables

meta data in this data set is ID(unique identifier), Title(reason of road strike), Location(the location of road,suburb and region), Created(the date which of event created), Coordination(longitude and latitude of event), Ended(time of the event is ended), Version(the version of event with updating the situation), Impact(the impact of event to the traffic).

## revelant spatial and temporal range the data covers

the spatial i chose is choosing the LGA which been affected by flooding and pick some suburbs, then search the event with radius of 50 KM. the coverage of data is mainly in the disaster stricken central areas, multiple suburbs has been chose in data selection to have better coverage. also, some LGA like Armidale which have no or few stricken events has not been used as a sample object. Time range here i have chose from 15 May 2025 to 17 Jul 2025 which three days before flooding and end by the date of flooding is almost over which reported by [NSW government](https://www.nsw.gov.au/emergency/floods/recover/flood-recovery-updates), this could capture data better and prevent data omissions caused by the spread of floods.

## judgement process of road closure

to filter the reason for road closure caused by natural disaster, i chose to use filter() to keep the reason of raod closure i need. in the population, i selected only 'flooding' and ' ADVERSE water over road', also, i made a second filer which keep only the events which made road closed. Flooding could be explained there has flood which caused by natural disaster and lead to road closure, adverse water over road has weaker relationship with disaster, the reason i kept this group of event is also caused by heavy rain but did not reached the standard of rain. these two reasons of road closure are all caused by natural disaster but have different degree.

## population and sample representativeness

the population is all road stricken caused by this flooding in the LGA list from 15 May 2025 to 17 Jul 2025. the sample here is representative, i selected through the flooding map and chose the central area of the flood as main collection area and supplemented by the surrounding area. the data is collected by searching with suburbs with radius of 50 km, LGA with has small area and densely distributed will have better coverage, thus the sample here is covered most of area of flooding and all of data after filtered are event which road closed due to natural disaster.

## limitation

the limitation of data is firstly the sample coverage, though the data chosen could cover most area of LGA, some data in urban area could be missed in data collection. secondly, while using this searching method with radius of 50 km will sometimes collect the data which not required for analytics: data of other LGA. thirdly is the searching method, to collect the traffic stricken by flooding, there should use the map of rainfall and conined with topographic map and traffic map, only search by suburb of LGA could collect the data required but could be inaccurate. furthermore, the data after tidy is only kept the fist version of some event, which could not find out how long is the road been closed.

# Resource& License

the resource of data are all from [Live Traffic NSW](https://www.livetraffic.com/), the license of this data is under the Creative Commons Attribution 4.0 Australia licence which allowed to share and adopt under the term of Attribution and no additional restrictions.

# citation

```{r}
citation("dplyr")
citation("readr")
citation("tidyverse")
```
